---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

The first step in this analysis is to read the input dataset into R (note: the activity.zip file was manually decompressed to create an "activity.csv" file).

```{r init,echo=FALSE}
setwd("c:/data/source/r/github/RepData_PeerAssessment1");
```

```{r load,echo=TRUE}
dt <- fread("activity.csv");

# Check that the data was loaded ok:
summary(dt);

```


## What is mean total number of steps taken per day?

```{r}
# Generate summary data.table of steps per date:
dtStepsByDate <- dt[,list(stepCount=sum(steps,na.rm=TRUE)),by=date];

# Histogram of total steps per day:
qplot( stepCount,
       data=dtStepsByDate,
       ylab = "Number of Days",
       xlab = "Number of Steps",
       binwidth = 500
     ) + theme_bw();

```

```{r}
# Calculate mean steps per day:
mean(dtStepsByDate$stepCount)

# Calculate median steps per day:
median(dtStepsByDate$stepCount)

```


## What is the average daily activity pattern?

```{r}
# calculate summary of mean steps per interval:
dtMeanStepsByInterval <- dt[,list(meanSteps=mean(steps,na.rm=TRUE)),by=interval]

# Plot time-series of average steps per interval:
ggplot(dtMeanStepsByInterval, aes(interval, meanSteps)) + geom_line();
```

```{r}
# Identify 5-min interval with highest step count on average:
dtMeanStepsByInterval[meanSteps == max(dtMeanStepsByInterval$meanSteps),];
```

## Imputing missing values

```{r}
# Determine number of rows with NAs:
nrow(dt[is.na(steps)]);
```

```{r}
# Make a copy of the data.table so that the fix isn't a	applied to the original data:
dtFix = copy(dt);

# Fill in gaps with values calculated in Mean Steps by Interval summmary data.table:
dtFix[is.na(steps)]$steps = 
  as.integer(
    ceiling(
      dtMeanStepsByInterval[interval == dt[is.na(steps)]$interval]$meanSteps
    )
  )

```

Now repeating steps but for fixed dataset:

```{r}
# Generate summary data.table of steps per date (fixed dataset):
dtfixStepsByDate <- dtFix[,list(stepCount=sum(steps,na.rm=TRUE)),by=date];

# Histogram of total steps per day (fixed dataset):
qplot( stepCount,
       data=dtfixStepsByDate,
       ylab = "Number of Days",
       xlab = "Number of Steps (fixed)",
       binwidth = 500
     ) + theme_bw();

```

```{r}
# Calculate mean steps per day (fixed dataset):
mean(dtfixStepsByDate$stepCount)

# Calculate median steps per day (fixed dataset):
median(dtfixStepsByDate$stepCount)

```

*Do these values differ from the estimates from the first part of the assignment?*  

**Yes**


*What is the impact of imputing missing data on the estimates of the total daily number of steps?*  

**The resulting mean / median are higher.**


## Are there differences in activity patterns between weekdays and weekends?

```{r}
# Add Weekend boolean factor to data.table:
dtFix1 <- transform(
            dtFix,
	    Weekend = substr(
	                weekdays(
			  strptime(dtFix$date,format="%Y-%m-%d")
			,TRUE)
	   	      ,1,1) == "S")

# Add "wd" factor with better naming for plotting purposes:
dtFix1$wd = factor(
              dtFix1$Weekend,
	      levels = c(TRUE,FALSE),
	      labels = c("weekend","weekday")
	    )

# Calculate mean grouped by interval and "wd" (week/weekend factor):
dtFixWDStepsByInterval <- dtFix1[,list(meanSteps=mean(steps)),by=c("interval","wd")]

# Generate plot:
xyplot(
    meanSteps ~ interval | wd,
    data = dtFixWDStepsByInterval,
    type = "l",
    layout = c(1,2)
);









